Demo arXiv template

### 2022-10-13 paper/paper.qmd
add changes/ todos after meeting

<!-- Index construction is a way to summarise complicated information (used in environmental data). **it basically says nothing** -->
<!-- The complexity of these information may involve a spatial distribution, a temporal scale that defines the frequency of the data, and a multivariate perspective that collects different climate variables. -->

<!-- These indices are customised to distinct local weather condition and historical data availability. -->

<!-- Various methods are proposed to extract multivariate information on the spatial coverage, across time. -->
<!-- Each individual index follows its own data pipeline and it can be difficult to evaluate how an index can be affected by tweaking parameters in a certain step, rearranging the order of steps, using to different method.  -->

<!-- Indices with readily available code to implement have more potential to be computed by researchers, i.e. the package `SPEI` [@SPEI] has implemented the SPI and SPEI index for calculating drought indices.  -->
<!-- [Another package in a different domain].   -->
<!-- However, even if most of the indices have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember, classify, and compare between. -->
<!-- There is a need for a framework to generally describe how indices can be constructed from spatio-temporal data. -->

<!-- This paper proposes such data pipeline for constructing indices. -->
<!-- The steps involved in the pipeline are general in nature and flexible to be adopted to most index construction for environmental data. -->

@@

@@

<!--   - decision making: need to consider multi-facet for allocating resources, [...] -->
<!--   - to include or constructed to quantify a concept of interest: some papers on constructing tourism based climate indices etc -->

<!-- Integrate multiple information to produce a single number summary (across time). -->

@@

The meaning of indices varies across disciplines and in this article, an index is defined as a tool to quantify a concept of interest that does not have a direct measure.
The concept of interest doesn't have a direct measure can because it is impractical to measure at the population level. 
For example, it would be nearly impossible to include all the available stocks in the market to characterise stock market behavior, so indices like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite select a representative set of stocks to measure the overall market behavior. 
Also belonging to this category are the economic indices like the Consumer Price Index, where price changes of a basket of items are weighted to measure inflation. 
The lack of direct measure could also because the concept itself is an unobservable human construction, rather than a physical quantity that can be measured.
Many natural hazard and social concepts falls into this category. 
This includes drought indices constructed from meteorological, agricultural, hydrological, and social-economic variables, e.g. Standardised Precipitation Index (SPI) [@mckee1993relationship] and Aggregated Drought Index (ADI) [@keyantash_aggregate_2004] among others.
Social development indices like Human Development Index [@hdi] and Global Liveability Index [@gli] measure various aspects of the quality of human capital and urban life.

@@

FYI:

  - Vegetation Condition Index  = (NDVI - NDVI_min)/ (NDVI_max - NDVI_min)
  - A measure is a variable, but is not an index. 
  - Not all multivariate quantities are indices (i.e. those remote sensor measures). It is possible that an index is constructed from a single variable (i.e. SPI).
  - Indices is not a single/ set of measures

**What is the problem with current workflow on index construction**

There has not yet been a standardised workflow on index construction. 
For simple indices, users may be able to work out the calculation from the equations or workflow provided, while for more complicated construction, software is not even available for open-source consumption.
Even when the software implementation is available, most indices are computed through a single index command, for example, the package `SPEI` provides two drought indices: SPI and SPEI through function `spi()` and `spei()`. 
This software design focuses on the name of the index and includes all the steps in a single index function. 
With this design, there is little room for modification. For example, fit a new distribution that is not provided in the initial code. 
Also, when scaled to a large number of indices, it increases the cognitive load of remembering hundreds of index names. 
A better approach is to modularise these steps and construct a pipeline for users to build indices through choosing different building blocks from each module.
With this design, users focus on the operations involved in the index construction.
If an operation from another index is useful, users can borrow it to modify an existing index.

@@

The article contributes a pipeline framework to construct general indices.

  - domain scientists who propose new indices and directly use indices in analysis. [use multiple indices to compare]
  - general scientists who uses indices as part of analysis, modelling process. Official indices are usually published in a public table with values calculated. Although methods of calculation are provided by the relevant index provider, raw data and scripts are usually not available for public consumption. The pipeline proposed in the paper opens up the possibility for general scientists to construct indices customised to their research purpose. 
  - decision makers, towards understand the 
  - statisticians, so as to [draw focus/ advocate] on statistical workflow and software design.

@@

@@

@@

A conversion of non-normal variables into (approximately) normal.

>>>>

<!-- Each individual index follows its own data pipeline and it can be difficult to evaluate how an index can be affected by tweaking parameters in a certain step, rearranging the order of steps, using to different method.  -->

<!-- However, even if most of the indices have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember, classify, and compare between. -->

@@

*incorporate the following in why using index: multiple pieces of information (variables) that need to be taken into account*

@@

@@

In this article, an index is defined as a tool to quantify a concept of interest that does not have a direct measure.
The concept of interest doesn't have a direct measure can because it is impractical to measure at the population level. 
For example, it would be nearly impossible to include all the available stocks in the market to characterise stock market behavior, so indices like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite select a representative set of stocks to measure the overall market behavior. 
Also belonging to this category are the economic indices like the Consumer Price Index, where price changes of a basket of items are weighted to measure inflation. 
The lack of direct measure could also because the concept itself is an unobservable human construction, rather than a physical quantity that can be measured.
Many natural hazard and social concepts falls into this category. 
This includes drought indices constructed from meteorological, agricultural, hydrological, and social-economic variables, e.g. Standardised Precipitation Index (SPI) [@mckee1993relationship] and Aggregated Drought Index (ADI) [@keyantash_aggregate_2004] among others.
Social development indices like Human Development Index [@hdi] and Global Liveability Index [@gli] measure various aspects of the quality of human capital and urban life.

*still need to tweak the tone a bit: "they are called index, they are not the index we will talk about"*

@@

<!-- FYI:  -->

<!--   - A measure is a variable, but is not an index.  -->
<!--   - Not all multivariate quantities are indices (i.e. those remote sensor measures). It is possible that an index is constructed from a single variable (i.e. SPI). -->
<!--   - Indices is not a single/ set of measures -->

**What is the problem with current index construction**

*indices constructed separately, but there are lots of are same [!], but hard to see from the current construction. "messy data are messy in there own way". Describe how they are same, so we can modular them*

There has not yet been a standardised workflow on index construction. 
For simple indices, users may be able to work out the calculation from the equations or workflow provided, while for more complicated construction, software is not even available for open-source consumption.
Even when the software implementation is available, most indices are computed through a single index command, [given examples but not too domain specific]. 
This software design focuses on the name of the index and includes all the steps in a single index function. 
With this design, there is little room for modification. For example, fit a new distribution that is not provided in the initial code. 
Also, when scaled to a large number of indices, it increases the cognitive load of remembering hundreds of index names. 
A better approach is to modularise these steps and construct a pipeline for users to build indices through choosing different building blocks from each module.
With this design, users focus on the operations involved in the index construction.
If an operation from another index is useful, users can borrow it to modify an existing index.

@@

*this work could be interested to scientists actively developing new indices*

  - domain scientists who propose new indices and directly use indices in analysis. [use multiple indices to compare]
  - general scientists who uses indices as part of analysis, modelling process. The pipeline proposed in the paper opens up the possibility for general scientists to construct indices customised to their research purpose. 
  - statisticians, so as to [draw focus/ advocate] on statistical workflow and software design.

**what can be done if people adopt this pipeline/ why it is beneficial?**

- transparency: official published, can decompose the indices, 
- allow for calculation on uncertainty with randomization, 
- compare indices,
- tweak/ swap building blocks, 
- create new indices based on combinations of existing steps.

@@

- easy to operate properly: should be designed that users know what is approriate to do
- promote good scientific methodology: should protect uses from doing stupid things,  doing right thing

good statistical practice

@@

Normalising is usually the last step

@@

change the shape of the variable

<!-- A conversion of non-normal variables into (approximately) normal. wrong! -->

### 2022-10-13 paper/paper.tex
add changes/ todos after meeting

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

@@

The meaning of indices varies across disciplines and in this article, an
index is defined as a tool to quantify a concept of interest that does
not have a direct measure. The concept of interest doesn't have a direct
measure can because it is impractical to measure at the population
level. For example, it would be nearly impossible to include all the
available stocks in the market to characterise stock market behavior, so
indices like Dow Jones Industrial Average, S\&P 500, and Nasdaq
Composite select a representative set of stocks to measure the overall
market behavior. Also belonging to this category are the economic
indices like the Consumer Price Index, where price changes of a basket
of items are weighted to measure inflation. The lack of direct measure
could also because the concept itself is an unobservable human
construction, rather than a physical quantity that can be measured. Many
natural hazard and social concepts falls into this category. This
includes drought indices constructed from meteorological, agricultural,
hydrological, and social-economic variables, e.g.~Standardised
Precipitation Index (SPI) (McKee et al. 1993) and Aggregated Drought
Index (ADI) (Keyantash and Dracup 2004) among others. Social development
indices like Human Development Index (United Nations Development
Programme 2022) and Global Liveability Index (Economist Intelligence
Unit 2019) measure various aspects of the quality of human capital and
urban life.

@@

FYI:

\begin{itemize}
\tightlist
\item
  Vegetation Condition Index = (NDVI - NDVI\_min)/ (NDVI\_max -
  NDVI\_min)
\item
  A measure is a variable, but is not an index.
\item
  Not all multivariate quantities are indices (i.e.~those remote sensor
  measures). It is possible that an index is constructed from a single
  variable (i.e.~SPI).
\item
  Indices is not a single/ set of measures
\end{itemize}

\textbf{What is the problem with current workflow on index construction}

There has not yet been a standardised workflow on index construction.
For simple indices, users may be able to work out the calculation from
the equations or workflow provided, while for more complicated
construction, software is not even available for open-source
consumption. Even when the software implementation is available, most
indices are computed through a single index command, for example, the
package \texttt{SPEI} provides two drought indices: SPI and SPEI through
function \texttt{spi()} and \texttt{spei()}. This software design
focuses on the name of the index and includes all the steps in a single
index function. With this design, there is little room for modification.
For example, fit a new distribution that is not provided in the initial
code. Also, when scaled to a large number of indices, it increases the
cognitive load of remembering hundreds of index names. A better approach
is to modularise these steps and construct a pipeline for users to build
indices through choosing different building blocks from each module.
With this design, users focus on the operations involved in the index
construction. If an operation from another index is useful, users can
borrow it to modify an existing index.

@@

The article contributes a pipeline framework to construct general
indices.

\begin{itemize}
\tightlist
\item
  domain scientists who propose new indices and directly use indices in
  analysis. {[}use multiple indices to compare{]}
\item
  general scientists who uses indices as part of analysis, modelling
  process. Official indices are usually published in a public table with
  values calculated. Although methods of calculation are provided by the
  relevant index provider, raw data and scripts are usually not
  available for public consumption. The pipeline proposed in the paper
  opens up the possibility for general scientists to construct indices
  customised to their research purpose.
\item
  decision makers, towards understand the
\item
  statisticians, so as to {[}draw focus/ advocate{]} on statistical
  workflow and software design.
\end{itemize}

@@

@@

A conversion of non-normal variables into (approximately) normal.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

\emph{incorporate the following in why using index: multiple pieces of
information (variables) that need to be taken into account}

@@

In this article, an index is defined as a tool to quantify a concept of
interest that does not have a direct measure. The concept of interest
doesn't have a direct measure can because it is impractical to measure
at the population level. For example, it would be nearly impossible to
include all the available stocks in the market to characterise stock
market behavior, so indices like Dow Jones Industrial Average, S\&P 500,
and Nasdaq Composite select a representative set of stocks to measure
the overall market behavior. Also belonging to this category are the
economic indices like the Consumer Price Index, where price changes of a
basket of items are weighted to measure inflation. The lack of direct
measure could also because the concept itself is an unobservable human
construction, rather than a physical quantity that can be measured. Many
natural hazard and social concepts falls into this category. This
includes drought indices constructed from meteorological, agricultural,
hydrological, and social-economic variables, e.g.~Standardised
Precipitation Index (SPI) (McKee et al. 1993) and Aggregated Drought
Index (ADI) (Keyantash and Dracup 2004) among others. Social development
indices like Human Development Index (United Nations Development
Programme 2022) and Global Liveability Index (Economist Intelligence
Unit 2019) measure various aspects of the quality of human capital and
urban life.

\emph{still need to tweak the tone a bit: ``they are called index, they
are not the index we will talk about''}

@@

\textbf{What is the problem with current index construction}

\emph{indices constructed separately, but there are lots of are same
{[}!{]}, but hard to see from the current construction. ``messy data are
messy in there own way''. Describe how they are same, so we can modular
them}

There has not yet been a standardised workflow on index construction.
For simple indices, users may be able to work out the calculation from
the equations or workflow provided, while for more complicated
construction, software is not even available for open-source
consumption. Even when the software implementation is available, most
indices are computed through a single index command, {[}given examples
but not too domain specific{]}. This software design focuses on the name
of the index and includes all the steps in a single index function. With
this design, there is little room for modification. For example, fit a
new distribution that is not provided in the initial code. Also, when
scaled to a large number of indices, it increases the cognitive load of
remembering hundreds of index names. A better approach is to modularise
these steps and construct a pipeline for users to build indices through
choosing different building blocks from each module. With this design,
users focus on the operations involved in the index construction. If an
operation from another index is useful, users can borrow it to modify an
existing index.

@@

\emph{this work could be interested to scientists actively developing
new indices}

\begin{itemize}
\tightlist
\item
  domain scientists who propose new indices and directly use indices in
  analysis. {[}use multiple indices to compare{]}
\item
  general scientists who uses indices as part of analysis, modelling
  process. The pipeline proposed in the paper opens up the possibility
  for general scientists to construct indices customised to their
  research purpose.
\item
  statisticians, so as to {[}draw focus/ advocate{]} on statistical
  workflow and software design.
\end{itemize}

\textbf{what can be done if people adopt this pipeline/ why it is
beneficial?}

\begin{itemize}
\tightlist
\item
  transparency: official published, can decompose the indices,
\item
  allow for calculation on uncertainty with randomization,
\item
  compare indices,
\item
  tweak/ swap building blocks,
\item
  create new indices based on combinations of existing steps.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  easy to operate properly: should be designed that users know what is
  approriate to do
\item
  promote good scientific methodology: should protect uses from doing
  stupid things, doing right thing
\end{itemize}

good statistical practice

@@

change the shape of the variable
